Exploring the Role of Small Molecules in Biological Systems Using Network Approaches

### 2015-03-22 paper.tex
Updated some formatting

While many network analysis packages in R focus on social science
applications, there are a plethora of R packages that focus on
biological networks. For example, the WCGNA package
\cite{Langfelder:2008cj} implements methods to construct gene networks
based on co-expression data. The Corbi package \cite{Huang:2013sy}
implements the CNetQ algorithm for network alignment, a task that is
relevant for predicting new protein-protein interactions. Given the
integrative nature of many biological problems that require the use of
multiple data types (e.g., genomic, proteomic, small molecule),
network approaches are commonly applied to investigating them. A
number of R packages support such integrative analyses - BioNet
\cite{Beisser:2010xp} is an R package that supports a variety of
integrative analyses, with a particular focus on identifying
functional modules in $k$-partite networks. NetWeAvers
\cite{McClellan:2013zf} is similar in nature, though it has
specialized support for mass spectrometry data. A number of packages
support statistical operations on graphs. For example,
\texttt{DEGraph} implements a method to test for differential
expression of gene networks \cite{Jacob:2012yg}. The \texttt{SANTA}
package \cite{Cornish:2014vs} implements methods to test the
association between a network and a phenotype. Several packages are
also available for enrichment analyses including \texttt{FGNet},
\texttt{miRNApath} and \texttt{viper}.

@@

A variety of algorithms related to repositioning and activity
predictions, based on drug-target networks have been implemented in
R. Ma and Zhao \cite{Ma:2012qy} implemented the FacPad method that
utilizes a Bayesian sparse factor model to identify pathways that are
perturbed upon drug treatments. The DrugComboRanker package
\cite{Huang:2014by} employed a community based partitioning to create
a drug meta-community network and a disease signaling meta-community
network. Drug targets were then mapped onto disease networks to allow
identification of synergistic pairs of drugs and complementary
mechanisms of action. In the tool called MANTRA (Mode of Action by
NeTwoRk Analysis) \cite{iorio2010discovery}, authors used network
communities to deduce mechanism of action of new molecules and suggest
repositioning of a well known molecule to a new therapeutic
arean. Finally the TIMMA-R package \cite{He:2015ly} combines binary
drug-target data and drug-sensitivity profiles to predict synergistic
pairs of anti-cancer drugs.

>>>>

While many network analysis packages in R focus on social science
applications, there are a plethora of R packages that focus on
biological networks. For example, the \texttt{WCGNA} package
\cite{Langfelder:2008cj} implements methods to construct gene networks
based on co-expression data. The \texttt{Corbi} package \cite{Huang:2013sy}
implements the CNetQ algorithm for network alignment, a task that is
relevant for predicting new protein-protein interactions. Given the
integrative nature of many biological problems that require the use of
multiple data types (e.g., genomic, proteomic, small molecule),
network approaches are commonly applied to investigating them. A
number of R packages support such integrative analyses - \texttt{BioNet}
\cite{Beisser:2010xp} is an R package that supports a variety of
integrative analyses, with a particular focus on identifying
functional modules in $k$-partite networks. \texttt{NetWeAvers}
\cite{McClellan:2013zf} is similar in nature, though it has
specialized support for mass spectrometry data. A number of packages
support statistical operations on graphs. For example,
\texttt{DEGraph} implements a method to test for differential
expression of gene networks \cite{Jacob:2012yg}. The \texttt{SANTA}
package \cite{Cornish:2014vs} implements methods to test the
association between a network and a phenotype. Several packages are
also available for enrichment analyses including \texttt{FGNet},
\texttt{miRNApath} and \texttt{viper}.

@@

A variety of algorithms related to repositioning and activity
predictions, based on drug-target networks have been implemented in
R. Ma and Zhao \cite{Ma:2012qy} implemented the \texttt{FacPad} method
that utilizes a Bayesian sparse factor model to identify pathways that
are perturbed upon drug treatments. The \texttt{DrugComboRanker}
package \cite{Huang:2014by} employed a community based partitioning to
create a drug meta-community network and a disease signaling
meta-community network. Drug targets were then mapped onto disease
networks to allow identification of synergistic pairs of drugs and
complementary mechanisms of action. In the tool called \texttt{MANTRA}
(Mode of Action by NeTwoRk Analysis) \cite{iorio2010discovery},
authors used network communities to deduce mechanism of action of new
molecules and suggest repositioning of a well known molecule to a new
therapeutic arean. Finally the \texttt{TIMMA-R} package
\cite{He:2015ly} combines binary drug-target data and drug-sensitivity
profiles to predict synergistic pairs of anti-cancer drugs.
